### 2018-11-26 manuscript/manuscript.Rmd
update manuscipt

---
title: "manuscript"
author: "Yuanning Li"
date: "10/17/2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
bibliography: manuscript.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Lamellibrachia genome
## Abstract

## Introduction
![](../figures/Picture1.png) 
**Fig 1.** *Lamellibrachia* from seep localities in Gulf of Mexico.

## Methods

### Biological materials.
Adult *Lamellibrachia* *luymesi* specimens were collected from from seep localities in the Mississippi Canyon at 754 m depth in Gulf of Mexico (N 28°11.58', W 89°47.94'), using the R/V Seward Johnson and Johnson Sea Link in October 2009. All samples were frozen at 80˚C following collection.

### Genome sequencing and assembly. 
Vestimentum tissue was dissected from one individual of worm, and high molecular weight genomic DNA was extracted using the the DNeasy Blood & Tissue Kit (Qiagen) according to the manufacturer’s protocols. Sequencing a total of six paired-end or mate-pair genomic DNA libraries with insert sizes ranging from 180 bp to 7 kb were performed by by The Genomic Services Lab at the Hudson Alpha Institute in Huntsville, Alabama on an Illumina HiSeq 2000 platform (see details in Table S1). Paired-end libraries (180 bp, 400 bp, 750 bp) were prepared using the 125 bp TrueSeq protocols, and mate-pair libraries (3-5 kbp, 5-7 kbp) were generated using the Illuomina Nextera Mate Pair Library Kit followed by size selection. Moreover, a 10X sequencing library was constructed using the 10X Chromium protocol (10X genomics) at the Hudson Alpha Institute. The finished library were sequenced on an Illumina HiSeqX platform, using paried 151 bp reads with a single 8 bp index read.

Our workflow of genome assembly was shown in Fig. S1. The paired-end and 10X raw reads were checked using FastQC v0.11.5 [@andrews2010fastqc] and quality filtered (Q score >30) using Trimmomatic v0.36 [@bolger2014trimmomatic]. The estimatation of genome size, level of heterozygosity and repeat contes of the Lamellibrachia genome was determined by analaysing the kmer histograms generated from the paired-end librries using Jellyfish v2.2.3 [@marccais2011fast] and GenomeScope [@vurture2017genomescope] (Fig. S2). The Mate-pair reads were trimmed and sorted using NxTrim v0.3.1 [@o2015nxtrim] which can recgonize and trim the artificial Nextera mate-pair circulation adapters. Only reads from category "mp" (true mate-pair reads) and "unkonwn" (mostly large insert size reads) were used for downstream scaffolding anlaysis. Reads from "pe" (paired-end reads) and "se" (single ends) categories were discarded.

Given that high heterozygosity of *Lamellibrachia* genome, all reads were assembled using Platanus v1.2.4 [@kajitani2014efficient] with a kmer size of 32. Scaffolding was conducted by mapping Illumina paired-end and mate-pair reads to contigs genrated by Platanus using SSPACE v3.0 [@boetzer2014sspace]. Gaps in the scaffolds were then filled with GapCloser v1.12 [@luo2012soapdenovo2]. Redundant allele scaffods were further remvoed using Redundans v0.13c with default settings [@pryszcz2016redundans].Our workflow of genome assembly of *Lamellibrachia* was shown in Fig. S2. Genome assembly quality was assessed using QUAST v3.1 [@gurevich2013quast]. Completeness of obtained genome was assessed using BUSCO v3[@waterhouse2017busco] with Metazoa_odb9 database (978 busco genes) .

## Results and Discussion

## Conclusion

## Ackowledgements

## Author contribution

\pagebreak

## Supplemental Information